\bf Approximating Likelihood Ratios with Calibrated Discriminative Classifiers

### 2016-02-10 jasa/paper-jasa.tex
Regenerate figures using same data for all classifiers

As a simple but illustrative example, let us first consider the approximation of
the likelihood log-ratio $\log \left( r(\mathbf{x};\theta_0,\theta_1) \right)$ between the mixtures
$p(\mathbf{x}|\theta_0)$ and $p(\mathbf{x}|\theta_1)$ illustrated in
Figure~\ref{fig:1} and defined as
\begin{align}
p(\mathbf{x}|\theta_0) &= (\frac{1}{2} - g) p_{c_0}(\mathbf{x}) + (\frac{1}{2} - g) p_{c_1}(\mathbf{x}) + g p_{c_2}(\mathbf{x}), \\
p(\mathbf{x}|\theta_1) &= \frac{1}{2} p_{c_0}(\mathbf{x}) + \frac{1}{2} p_{c_1}(\mathbf{x}),
\end{align}
where $p_{c_0} := {\cal N}(\mu=-2, \sigma^2=0.5625)$, $p_{c_1} := {\cal N}(\mu=0, \sigma^2=4)$,
$p_{c_2} := {\cal N}(\mu=1, \sigma^2=0.25)$ and where $g$ is a parameter of $p(\mathbf{x}|\theta_0)$
set to $0.05$.

@@

Let us now consider the composite hypothesis testing setting, by comparing
$H_0: g = 0$ versus the alternative $H_1: g > 0$. That is, we want to evaluate
\begin{equation}
\Lambda({\cal D}; \Theta_0 = \{0\}, \Theta=(0, \frac{1}{2}]) =  \frac{ p({\cal D} | g = 0) }{ \sup_{\theta \in \Theta} p({\cal D} | g=\theta)} \; .
\end{equation}
Following the approach outlined in Section~\ref{sec:generalized-likelihood-ratio},
computing the generalized likelihood ratio statistic requires finding
the maximum likelihood estimator for the denominator, which can be carried out using
approximated likelihood ratios. Reusing the decomposed neural network-based
approximation (or equivalently, using a parameterized classifier) and minimizing the mean negative log-ratio of some observed data
generated for $g=0.05$,
we successfully find as estimator $\hat g=0.049$, as shown in Figure~\ref{fig:3}.
From there, the generalized likelihood ratio can be approximated as in the simple
hypothesis setting.

@@

\glnote{Smoothly interpolate between histogram bins, in order to get a smooth ratio as well. Worth mentioning in the paper!}
\glnote{Discuss diagnostics?}
\glnote{Better highlight advantages of the method?}
\glnote{Conduct more thorough and convincing experiments?}

>>>>

As a simple but illustrative example, let us first consider the approximation of
the likelihood log-ratio $\log \left( r(\mathbf{x};\theta_0,\theta_1) \right)$ between the mixtures
$p(\mathbf{x}|\theta_0)$ and $p(\mathbf{x}|\theta_1)$ illustrated in
Figure~\ref{fig:1} and defined as
\begin{align}
p(\mathbf{x}|\theta_0) &= (\frac{1}{2} - g) p_{c_0}(\mathbf{x}) + (\frac{1}{2} - g) p_{c_1}(\mathbf{x}) + g p_{c_2}(\mathbf{x}), \\
p(\mathbf{x}|\theta_1) &= \frac{1}{2} p_{c_0}(\mathbf{x}) + \frac{1}{2} p_{c_1}(\mathbf{x}),
\end{align}
where $p_{c_0} := {\cal N}(\mu=-2, \sigma^2=0.5625)$, $p_{c_1} := {\cal N}(\mu=0, \sigma^2=4)$,
$p_{c_2} := {\cal N}(\mu=1, \sigma^2=0.25)$ and where $g$ is a parameter of $p(\mathbf{x}|\theta_0)$
set to $0.05$. Samples drawn from $p(\mathbf{x}|\theta_0)$ are shown in the
figure and will be used later for inference.

@@

Let us now consider the composite hypothesis testing setting, by comparing
$H_0: g = 0$ versus the alternative $H_1: g > 0$. That is, we want to evaluate
\begin{equation}
\Lambda({\cal D}; \Theta_0 = \{0\}, \Theta=(0, \frac{1}{2}]) =  \frac{ p({\cal D} | g = 0) }{ \sup_{\theta \in \Theta} p({\cal D} | g=\theta)} \; .
\end{equation}
Following the approach outlined in Section~\ref{sec:generalized-likelihood-ratio},
computing the generalized likelihood ratio statistic requires finding
the maximum likelihood estimator for the denominator, which can be carried out using
approximated likelihood ratios. Reusing the decomposed neural network-based
approximation (or equivalently, using a parameterized classifier) and minimizing the mean negative log-ratio of some observed data
generated for $g=0.05$ (as illustrated in Figure~\ref{fig:1}),
we successfully find as estimator $\hat g=0.049$, as shown in Figure~\ref{fig:3}.
From there, the generalized likelihood ratio can be approximated as in the simple
hypothesis setting.

@@

\glnote{Smoothly interpolate between histogram bins, in order to get a smooth ratio as well?}
\glnote{Discuss diagnostics?}
\glnote{Better highlight advantages of the method?}
\glnote{Conduct more thorough and convincing experiments?}
